### 2015-10-21 proposal/narrative/data-desc.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

\subsubsection{Description of Data}
In order for the middleware and storage layer to understand data,
I/O intentions, and act accordingly, describing data in a way that can
be understood by the system is needed. In this project, this is achieved 
by managing data as \textit{objects}.
An object is the smallest unit of data that consists of raw data and metadata,
and a user space variable, such as \textit{temperature}, may consist of a collection
of objects possibly with different accuracy or resolution.
A key advantage of representing data as \textit{objects}, rather than 
bytes, is that it allows data semantics, user intentions, QoS requirements, and
the relationship between user data to be readily captured and embedded with the 
raw data. As an example, retrieving a 3D field data generated from a simulation and then visualizing it
requires the coordinates and connectivity variable 1) to be accessed together and 2) with low
latency in order to achieve a good user experience. The description of data, 
as enabled by managing data as objects, bridges 
the semantics gap between applictions, middleware, and storage systems,
and allows the system to understand user-level data, execute QoS requirements and policies, and
optimize application and system performance.

>>>>

\subsubsection{Description of Data}
In order for the middleware and storage layer to understand data,
I/O intentions, and act accordingly, describing data in a way that can
be understood by the system is needed. In this project, this is achieved 
by managing data as \textit{objects}.
An object is the smallest unit of data that consists of raw data and metadata,
and a user space variable, such as \textit{temperature}, may consist of a collection
of objects possibly with different accuracy or resolution.
A key advantage of representing data as \textit{objects}, rather than 
bytes, is that it allows data semantics, user intentions, QoS requirements, and
the relationship between user data to be readily captured and embedded with the 
raw data. As an example, retrieving a 3D field data generated from a simulation and then visualizing it
requires the coordinates and connectivity variable (1) to be accessed together and (2) with low
latency in order to achieve a good user experience. The description of data, 
as enabled by managing data as objects, bridges 
the semantics gap between applictions, middleware, and storage systems,
and allows the system to understand user-level data, execute QoS requirements and policies, and
optimize application and system performance.

### 2015-10-21 proposal/narrative/data-reduction.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

To refactor data effectively and efficiently, it is important to understand
the critical point when the time and resources required to
identify and perform the ``best methods'' outweighs the intended gains. Another
critical research question concerns the quantification and control of the loss
of information resulting from refactoring the data and using a reduced dataset. 
A basic issue allied to refactoring is to understand how much information is
actually present in a dataset and whether a refactoring based on a reduced
order representation might prove effective. It is useful to classify scientific
data into two basic categories: structured data (in the sense that it satisfies
a known or relatively simple model), or unstructured data (in which information
content follows no obvious or explicit model).  Although scientific data
generally contains random components (due to finite precision and measurement and
calibration effects), useful scientific data is never purely random.
Broadly speaking, the path from data to knowledge consists of extracting
underlying models or patterns from the datasets and interpreting the resulting
models.

>>>>

To refactor data effectively and efficiently, it is important to understand
the critical point when the time and resources required to
identify and perform the ``best methods'' outweighs the intended gains. Another
critical research question concerns the quantification and control of the loss
of information resulting from refactoring the data and using a reduced dataset. 
A basic issue allied to refactoring is to understand how much information is
actually present in a dataset and whether a refactoring based on a reduced
order representation might prove effective. It is useful to classify scientific
data into two basic categories: regular data (in the sense that it satisfies
a known or relatively simple model), or irregular data (in which information
content follows no obvious or explicit model).  Although scientific data
generally contains random components (due to finite precision and measurement and
calibration effects), useful scientific data is never purely random.
Broadly speaking, the path from data to knowledge consists of extracting
underlying models or patterns from the datasets and interpreting the resulting
models.

### 2015-10-21 proposal/narrative/eviction-metrics.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

\paragraph{Approach:} This spectrum enables interesting strategies
to manage space pressure on high tiers: instead of just evicting
pieces of data (blocks, pages, objects, files), one can free space
by ``thinning'' data so that accesses within more limited views or
lower resolutions can still occur without misses while requests
outside of views or higher resolutions can leverage the information
stored in auxiliary data. Conversely, data on a high tier can be
``enriched'' (e.g. turned into a higher resolution) if there is
value in doing so.

>>>>

\paragraph{Approach:} This spectrum enables interesting strategies
to manage space pressure on high tiers: instead of just evicting
pieces of data (blocks, pages, objects, files), one can free space
by ``thinning'' data so that accesses within more limited views or
lower resolutions can still occur without misses while requests
accessing outside of views or higher resolutions can leverage the information
stored in auxiliary data. Conversely, data on a high tier can be
``enriched'' (e.g. turned into a higher resolution) if there is
value in doing so.

### 2015-10-21 proposal/narrative/intro.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

The motivating use case is a series of simulations of the
ITER~\cite{dietz1996iter} fusion experiments using the
XGC1~\cite{chang2006integrated} application. XGC1 is one of the largest
applications used at the Department of Energy's (DOE's) Leadership Class
Facilities, with an allocation of over 300M hours. 
% % simulations
% % simulations planne
% %that C. S. Chang (PPPL), 
% one of the largest users of Leadership Class Facilities (LCFs) (over 300M hours at DOE facilities), planned to launch.
These simulations were expected to produce 100~PB of data over 10 days of
total runtime on the Titan system at the Oak Ridge Leadership Computing
Facility (OLCF), and required a team of experts including
the user group and I/O and storage personnel to help ensure that the maximum
amount of information would be saved. Due to physical resource limitations,
the simulation was immediately restricted to write about 10~PB. However,
when the time to write and read, together with the financial cost to archive
this volume of data was fully explored, the data size goal was further
reduced to only 5~PB over the 10 day run.

@@

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\caption {Relevance to SSIO}
\label{tab:call}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant SIRIUS work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S\ref{sec:managing-data-life}, \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of middleware with the storage system & Manage data placement and migration between the application, middleware and storage system using an autonomic cross-layer approach & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline
\end{tabular}
\end{center}
\vskip -0.5cm
\end{table}

@@

One of our research objectives is to understand how the
storage system can interact with the middleware   so
that the data with the highest utility is kept on the fastest layers of the
storage system for as long as specified. We need to maintain 
QoS and therefore, be able to provide fuzzy estimates of the time it will take to write/read data from
the different layers to inform any migration/eviction decisions, and it
is also critical that the storage system can manage (discover and retrieve) the data across all of the
layers automatically, without user intervention. 
Our approach here is that the storage system needs to define bounds of time estimates
(min time, max time), and as our system evolves, we want to allow plug-ins which can help
reduce the range of the bounds.
Overall, our goal is to enable predictable performance acorss the storage system, and we will develop mechanisms to enable this
(\S\ref{sec:naming-discovery}).

@@

From the reading perspective, the data must be accessible in a fast and
predictable manner for all read access patterns, with different access
patterns having their own requirements for latency, throughput and
consistency. Here the user must be able to describe the read patterns and
query for performance estimations (\S\ref{sec:data-description}), be able
to quickly access the most important data (\S\ref{sec:data-refactor}) and recieve
performance gurantees (\S\ref{sec:qos}) on the time required to complete
the read request. The challenge here is to address the variety of read
patterns that drive knowledge discover.

@@

Our research objective here is to explore the interface that reading code
can utilize to discover the important properites of data, the expectations
about the time to access the data, and make decisions on the quality of data
required to complete the task.

@@

%negotation between these layers will be a fundamental service which we will

@@

%hardware. The storage layer will be partitioned into multiple heterogenous

>>>>

The motivating use case is a series of simulations of the
ITER~\cite{dietz1996iter} fusion experiments using the
XGC1~\cite{chang2006integrated} application. XGC1 is one of the largest
applications used at the Department of Energy's (DOE's) Leadership Class
Facilities, with an allocation of over 300M hours. 
% % simulations
% % simulations plane
% %that C. S. Chang (PPPL), 
% one of the largest users of Leadership Class Facilities (LCFs) (over 300M hours at DOE facilities), planned to launch.
These simulations were expected to produce 100~PB of data over 10 days of
total runtime on the Titan system at the Oak Ridge Leadership Computing
Facility (OLCF), and required a team of experts including
the user group and I/O and storage personnel to help ensure that the maximum
amount of information would be saved. Due to physical resource limitations,
the simulation was immediately restricted to write about 10~PB. However,
when the time to write and read, together with the financial cost to archive
this volume of data was fully explored, the data size goal was further
reduced to only 5~PB over the 10 day run.

@@

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\caption {Relevance to SSIO}
\label{tab:call}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant SIRIUS work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S\ref{sec:managing-data-life}, \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of middleware with the storage system & Manage data placement and migration between the application, middleware and storage system using an autonomic cross-layer approach & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow C/R data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline
\end{tabular}
\end{center}
\vskip -0.5cm
\end{table}

@@

One of our research objectives is to understand how the
storage system can interact with the middleware   so
that the data with the highest utility is kept on the fastest layers of the
storage system for as long as specified. We need to maintain 
QoS and therefore, be able to provide fuzzy estimates of the time it will take to write/read data from
the different layers to inform any migration/eviction decisions, and it
is also critical that the storage system can manage (discover and retrieve) the data across all of the
layers automatically, without user intervention. 
Our approach here is that the storage system needs to define bounds of time estimates
(min time, max time), and as our system evolves, we want to allow plug-ins which can help
reduce the range of the bounds.
Overall, our goal is to enable predictable performance across the storage system, and we will develop mechanisms to enable this
(\S\ref{sec:naming-discovery}).

@@

From the reading perspective, the data must be accessible in a fast and
predictable manner for all read access patterns, with different access
patterns having their own requirements for latency, throughput and
consistency. Here the user must be able to describe the read patterns and
query for performance estimations (\S\ref{sec:data-description}), be able
to quickly access the most important data (\S\ref{sec:data-refactor}) and receive
performance guarantees (\S\ref{sec:qos}) on the time required to complete
the read request. The challenge here is to address the variety of read
patterns that drive knowledge discover.

@@

Our research objective here is to explore the interface that reading code
can utilize to discover the important properties of data, the expectations
about the time to access the data, and make decisions on the quality of data
required to complete the task.

@@

%negotiation between these layers will be a fundamental service which we will

@@

%hardware. The storage layer will be partitioned into multiple heterogeneous

### 2015-10-21 proposal/narrative/migration.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

One important research question in this proposal is whether and how a liberated
Hierarchical Storage Manager designed for online use, such as Sirocco, might be
employed in supporting the higher level concepts discussed previously.
Conceptually, in this effort, data is accessed by the middleware IO libraries
at various tiers in the lower-level storage hierarchy. While Sirocco does
autonomously group the distributed set of media into like groups and manage
movement of data between these groups the various motivations to do so include
both the traditional ones found in classical HSM (Hierarchical Storage
Management) as well as the novel.

>>>>

One important research question in this proposal is whether and how a liberated
Hierarchical Storage Manager designed for online use, such as Sirocco, might be
employed in supporting the higher level concepts discussed previously.
Conceptually, in this effort, data is accessed by the middleware IO libraries
at various tiers in the lower-level storage hierarchy. While Sirocco does
autonomously group the distributed set of media into groups and manage
movement of data between these groups the various motivations to do so include
both the traditional ones found in classical HSM (Hierarchical Storage
Management) as well as the novel.

### 2015-10-21 proposal/narrative/proposed-work.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

Additionally we recognize that HPC applications exhibit a set
of common I/O patterns \cite{lofstead2011six,polte2009and,tian2011edo,tian2012system}.
Throughout our project we will continuously focus on creating research artifacts 
which can optimize these I/O patterns on current and
exascale systems. One area of focus is Checkpoint/Restart (C/R), leveraging NVRAM to accelerate
checkpoint and data retrieval. 
The migration and purging of this will be optimized by SIRUS using the utility function of 
the C/R data so that this can be stored for short periods of time, and then purged.

>>>>

Additionally we recognize that HPC applications exhibit a set
of common I/O patterns \cite{lofstead2011six,polte2009and,tian2011edo,tian2012system}.
Throughout our project we will continuously focus on creating research artifacts 
which can optimize these I/O patterns on current and
exascale systems. One area of focus is Checkpoint/Restart (C/R), leveraging NVRAM to accelerate
checkpoint and data retrieval. 
The migration and purging of this will be optimized by SIRIUS using the utility function of 
the C/R data so that this can be stored for short periods of time, and then purged.

### 2015-10-21 proposal/participant-table.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

% Generated by bin/config-participant-table on Fri, Jul 10, 2015  1:33:09 PM
\newlength{\longestpientry}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}}}
\begin{tabularx}{\textwidth}{l>{\raggedright\arraybackslash}X}
\emph{Institution} & \makebox[\longestpientry][l]{\emph{Principal Investigator (Email)}} \quad \emph{Additional Senior Personnel}
\\
\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Mark Ainsworth, Qing Liu, Kimmy Mu, Feiyi Wang\\
\textbf{\rutgers} & \makebox[\longestpientry][l]{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}\hspace*{1em}\\
\textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\\
\end{tabularx}

>>>>

% Generated by bin/config-participant-table on Fri, Jul 10, 2015  1:33:09 PM
\newlength{\longestpientry}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}}}
\begin{tabularx}{\textwidth}{l>{\raggedright\arraybackslash}X}
\emph{Institution} & \makebox[\longestpientry][l]{\emph{Principal Investigator (Email)}} \quad \emph{Additional Senior Personnel}
\\
\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Mark Ainsworth,\\
\textbf{} & \makebox[\longestpientry][l]{\textbf{ }\hspace*{1in}}       Qing Liu, Kimmy Mu, Feiyi Wang\\
\textbf{\rutgers} & \makebox[\longestpientry][l]{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}\hspace*{1em}\\
\textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\\
\end{tabularx}
